---
title: "Final Project"
author: "Qingya Yang"
format:
  html:
    toc: true
    toc-depth: 3
    theme: flatly
    embed-resources: true
execute:
  warning: false
  message: false
  echo: true
editor: visual
---

```{r}
library(readxl)
library(dplyr)
library(purrr)
library(stringr)
library(janitor)
library(tibble)
library(tidyr)
library(haven)
library(ranger)
library(glmnet)
library(xgboost)
library(Metrics)
library(knitr)
```

# Part 1: Data Cleaning

## 1.1 Target Variables: Annual change in log carbon intensity

```{r}
### Calculating Scope 1 Carbon Emissions ###

dir_path <- "D:/yqy/硕士-mpp/第四学期/ds&climate change/project/data/target_variable"

# List all Emission Inventories Excel files in the folder
files <- list.files(
  dir_path,
  pattern = "Emission_Inventories.*\\.xlsx$",
  full.names = TRUE,
  ignore.case = TRUE
)
cat("Files found:", length(files), "\n")

# Extract province name from sheet name by removing trailing 4-digit year
prov_from_sheet <- function(x) {
  x <- gsub("\\s+", "", x)
  sub("\\d{4}$", "", x)
}

# For each file, iterate over province sheets and extract Scope 1 Total Emissions
extract_scope1 <- function(file) {
  year <- as.integer(str_extract(basename(file), "\\d{4}"))
  sheets <- excel_sheets(file)
  
  # Skip the NOTE sheet (metadata only)
  sheets <- sheets[sheets != "NOTE"]
  
  map_dfr(sheets, function(sh) {
    tryCatch({
      df <- read_excel(file, sheet = sh, col_names = TRUE)
      df <- clean_names(df)
      
      # Ensure first column is named emission_inventory
      if (!"emission_inventory" %in% names(df)) names(df)[1] <- "emission_inventory"
      
      # Find the TotalEmissions row (strip spaces and lowercase for robust matching)
      row_idx <- which(
        tolower(gsub("\\s+", "", as.character(df$emission_inventory))) == "totalemissions"
      )
      
      # Find the Scope_1_Total column; fall back to regex if exact match fails
      cand <- which(tolower(names(df)) == "scope_1_total")
      if (length(cand) == 0)
        cand <- grep("scope.*1.*total", tolower(names(df)), perl = TRUE)
      
      # Skip sheet if either the row or column cannot be located
      if (length(cand) == 0 || length(row_idx) == 0) return(NULL)
      
      tibble(
        province     = prov_from_sheet(sh),
        year         = year,
        scope1_total = suppressWarnings(as.numeric(df[[cand[1]]][row_idx[1]]))
      )
    }, error = function(e) {
      message("Error in sheet ", sh, ": ", e$message)
      NULL
    })
  })
}

# Apply extraction across all files and combine into one data frame
final_scope1 <- map_dfr(files, extract_scope1) %>%
  filter(!is.na(scope1_total)) %>%
  arrange(province, year)

# Save output to CSV
out_file <- file.path(dir_path, "Scope1_Total_Emissions_2008_2022.csv")
write.csv(final_scope1, out_file, row.names = FALSE)

# Preview results
print(head(final_scope1, 10))

# Check number of provinces per year (should be ~30 each year)
final_scope1 %>%
  count(year) %>%
  arrange(year) %>%
  print(n = 50)

### Carbon Intensity Calculation ###
emissions <- final_scope1 

# Load Provincial GDP
gdp_path <- "D:/yqy/硕士-mpp/第四学期/ds&climate change/project/data/economic_structure/provincial_gdp_2008_2021.xlsx"

gdp_wide <- read_excel(gdp_path, skip = 2) 

# Reshape from wide to long (province, year, gdp)
gdp_long <- gdp_wide %>%
  rename(province = 1) %>%                   
  filter(!is.na(province)) %>%               
  pivot_longer(
    cols      = -province,
    names_to  = "year",
    values_to = "gdp_100m_yuan"              # Unit: 100 million yuan
  ) %>%
  mutate(year = as.integer(year))

# Harmonize province names between the two datasets
# Check for mismatches before joining
emissions_provs <- sort(unique(emissions$province))
gdp_provs       <- sort(unique(gdp_long$province))

cat("Provinces only in emissions:\n")
print(setdiff(emissions_provs, gdp_provs))
cat("Provinces only in GDP:\n")
print(setdiff(gdp_provs, emissions_provs))

# Fix the known mismatch
emissions <- emissions %>%
  mutate(province = recode(province, "InnerMongolia" = "Inner Mongolia"))

# Merge and calculate carbon intensity
carbon_intensity <- emissions %>%
  left_join(gdp_long, by = c("province", "year")) %>%
  filter(!is.na(gdp_100m_yuan)) %>%
  mutate(
    # Carbon intensity: Mt CO2 per 100 million yuan GDP
    ci       = scope1_total / gdp_100m_yuan,
    
    # Log carbon intensity
    ln_ci    = log(ci),
    
    # Annual change in log carbon intensity (target variable)
    d_ln_ci  = ln_ci - lag(ln_ci, 1)
  ) %>%
  
  # lag() must be applied within each province
  group_by(province) %>%
  arrange(year, .by_group = TRUE) %>%
  mutate(d_ln_ci = ln_ci - lag(ln_ci, 1)) %>%
  ungroup()

print(head(carbon_intensity, 10))

# Check coverage
cat("\nRows:", nrow(carbon_intensity), "\n")
carbon_intensity %>% count(year) %>% arrange(year) %>% print(n = 20)

# Save
econ_path <- "D:/yqy/硕士-mpp/第四学期/ds&climate change/project/data/target_variable"
out_ci <- file.path(econ_path, "carbon_intensity_2008_2021.csv")
write.csv(carbon_intensity, out_ci, row.names = FALSE)
cat("Saved to:", out_ci, "\n")
```

## 1.2 Control Variables

```{r}
### GDP per Capita Calculation ###
econ_path <- "D:/yqy/硕士-mpp/第四学期/ds&climate change/project/data/economic_structure"

# Load GDP data
gdp_wide <- read_excel(
  file.path(econ_path, "provincial_gdp_2008_2021.xlsx"),
  skip = 2  
)

gdp_long <- gdp_wide %>%
  rename(province = 1) %>%
  filter(!is.na(province), province != "Province") %>%  
  pivot_longer(
    cols      = -province,
    names_to  = "year",
    values_to = "gdp_100m_yuan"
  ) %>%
  mutate(year = as.integer(year))

# Load Population data (unit: 10,000 persons) 
province_order <- c(
  "Beijing", "Tianjin", "Hebei", "Shanxi", "Inner Mongolia",
  "Liaoning", "Jilin", "Heilongjiang",
  "Shanghai", "Jiangsu", "Zhejiang", "Anhui", "Fujian", "Jiangxi", "Shandong",
  "Henan", "Hubei", "Hunan", "Guangdong", "Guangxi", "Hainan",
  "Chongqing", "Sichuan", "Guizhou", "Yunnan", "Tibet",
  "Shaanxi", "Gansu", "Qinghai", "Ningxia", "Xinjiang"
)

pop_wide <- read_xls(
  file.path(econ_path, "provincial_population_2008_2021.xls"),
  skip = 0   # Keep row 0 as header (Region, 2021, 2020, ..., 2008)
)

pop_long <- pop_wide %>%
  select(-1) %>%                            # Drop the "Region" placeholder column
  mutate(province = province_order) %>%     # Assign province names by row position
  pivot_longer(
    cols      = -province,
    names_to  = "year",
    values_to = "population_10k"
  ) %>%
  mutate(year = as.integer(year))

# Calculate GDP per capita and log GDP per capita
gdp_percap <- gdp_long %>%
  left_join(pop_long, by = c("province", "year")) %>%
  filter(!is.na(population_10k), province != "Tibet") %>% 
  mutate(
    # GDP per capita in yuan: (100m yuan × 1e8) / (10k persons × 1e4)
    gdp_percap    = gdp_100m_yuan * 10000 / population_10k,
    ln_gdp_percap = log(gdp_percap)
  ) %>%
  arrange(province, year)

print(head(gdp_percap, 10))
cat("\nRows:", nrow(gdp_percap), "\n")  

out_gdp <- file.path(econ_path, "gdp_percap_2008_2021.csv")
write.csv(gdp_percap, out_gdp, row.names = FALSE)
cat("Saved to:", out_gdp, "\n")
```

# Part 2: Merge All Data and Run Models

## 2.1 Data Merging

```{r}
# Define paths
econ_path   <- "D:/yqy/硕士-mpp/第四学期/ds&climate change/project/data/economic_structure"
energy_path <- "D:/yqy/硕士-mpp/第四学期/ds&climate change/project/data/energy_structure"
policy_path <- "D:/yqy/硕士-mpp/第四学期/ds&climate change/project/data/policy"
target_path <- "D:/yqy/硕士-mpp/第四学期/ds&climate change/project/data/target_variable"

# Load each dataset
# Target variable
ci <- read.csv(file.path(target_path, "carbon_intensity_2008_2021.csv")) %>%
  select(province, year, ln_ci, d_ln_ci)

# GDP per capita
gdp_pc <- read.csv(file.path(econ_path, "gdp_percap_2008_2021.csv")) %>%
  select(province, year, ln_gdp_percap)

# Urban population rate
urban <- read_excel(file.path(econ_path, "urban_population_rate.xlsx"), skip = 2) %>%
  rename(province = 1) %>%
  filter(!is.na(province)) %>%
  pivot_longer(-province, names_to = "year", values_to = "urban_rate") %>%
  mutate(year = as.integer(year))

# Secondary industry share
sh_sec <- read_excel(
  file.path(econ_path, "industry_shares_2008_2021.xlsx"),
  sheet = "Secondary Industry Share", skip = 2
) %>%
  rename(province = 1) %>%
  filter(!is.na(province)) %>%
  pivot_longer(-province, names_to = "year", values_to = "sh_secondary") %>%
  mutate(year = as.integer(year))

# Tertiary industry share
sh_ter <- read_excel(
  file.path(econ_path, "industry_shares_2008_2021.xlsx"),
  sheet = "Tertiary Industry Share", skip = 2
) %>%
  rename(province = 1) %>%
  filter(!is.na(province)) %>%
  pivot_longer(-province, names_to = "year", values_to = "sh_tertiary") %>%
  mutate(year = as.integer(year))

# Coal share
coal <- read.csv(file.path(energy_path, "coal_share.csv")) %>%
  select(province, year, coal_share) %>%
  mutate(province = recode(province, "Inner mongolia" = "Inner Mongolia"))

# Policy intensity — filter to 2008-2021, drop Xizang
policy <- read_dta(file.path(policy_path, "Provincial_Policy_Intensity.dta")) %>%
  filter(year >= 2008, year <= 2021, province != "Xizang") %>%
  mutate(year = as.integer(year))

# Merge all into one panel
panel <- ci %>%
  left_join(gdp_pc,  by = c("province", "year")) %>%
  left_join(urban,   by = c("province", "year")) %>%
  left_join(sh_sec,  by = c("province", "year")) %>%
  left_join(sh_ter,  by = c("province", "year")) %>%
  left_join(coal,    by = c("province", "year")) %>%
  left_join(policy,  by = c("province", "year"))

cat("Panel rows after merge:", nrow(panel), "\n")
print(head(panel, 5))
```

## 2.2 Feature Engineering and Train/Test Split

```{r}
# Construct lagged features (t-1 predicts t)
features <- c("ln_gdp_percap", "urban_rate", "sh_secondary",
              "sh_tertiary", "coal_share", "PI_provincial", "ln_ci")

panel_lagged <- panel %>%
  group_by(province) %>%
  arrange(year, .by_group = TRUE) %>%
  mutate(across(all_of(features), ~ lag(., 1), .names = "{.col}_lag")) %>%
  ungroup() %>%
  filter(!is.na(d_ln_ci), !is.na(ln_gdp_percap_lag))

cat("Rows after lagging (model-ready):", nrow(panel_lagged), "\n")

lag_features <- paste0(features, "_lag")

# Train (2009-2017) / Test (2018-2021) split
train <- panel_lagged %>% filter(year <= 2017)
test  <- panel_lagged %>% filter(year >= 2018)

cat("Train rows:", nrow(train), " | Test rows:", nrow(test), "\n")

X_train <- as.matrix(train[, lag_features])
y_train <- train$d_ln_ci
X_test  <- as.matrix(test[, lag_features])
y_test  <- test$d_ln_ci
```

## 2.3 Model 0: Naïve Persistence Baseline

```{r}
# Predict next year's change using last year's log carbon intensity change
# i.e., assume d_ln_ci(t) ≈ d_ln_ci(t-1)
t_naive <- system.time({
  naive_pred <- panel_lagged %>%
    group_by(province) %>%
    arrange(year, .by_group = TRUE) %>%
    mutate(naive_pred = lag(d_ln_ci, 1)) %>%
    ungroup() %>%
    filter(year >= 2018, !is.na(naive_pred))
})["elapsed"]

# Align y_test to matched rows (some rows may drop due to missing lag)
naive_actual <- naive_pred$d_ln_ci
naive_hat    <- naive_pred$naive_pred

cat("\n--- Naïve Persistence Baseline ---\n")
cat("MAE: ",  round(mae(naive_actual,  naive_hat), 4), "\n")
cat("RMSE:", round(rmse(naive_actual, naive_hat), 4), "\n")
cat("R²:  ",  round(cor(naive_actual,  naive_hat)^2, 4), "\n")
```

## 2.4 Model 1: OLS

```{r}
train_df <- data.frame(d_ln_ci = y_train, X_train)
ols_formula <- as.formula(paste("d_ln_ci ~", paste(lag_features, collapse = " + ")))
ols_model <- lm(ols_formula, data = train_df)

t_ols <- system.time({
  ols_pred <- predict(ols_model, newdata = as.data.frame(X_test))
})["elapsed"]

cat("\n--- OLS Results ---\n")
cat("MAE: ",  round(mae(y_test, ols_pred), 4), "\n")
cat("RMSE:", round(rmse(y_test, ols_pred), 4), "\n")
cat("R²:  ",  round(cor(y_test, ols_pred)^2, 4), "\n")
```

## 2.5 Model 2: Ridge Regression

```{r}
# Use cross-validation to select optimal lambda
t_ridge <- system.time({
  set.seed(42)
  ridge_cv     <- cv.glmnet(X_train, y_train, alpha = 0, nfolds = 5)
  ridge_lambda <- ridge_cv$lambda.min
  ridge_model  <- glmnet(X_train, y_train, alpha = 0, lambda = ridge_lambda)
})["elapsed"]

ridge_pred <- as.vector(predict(ridge_model, newx = X_test, s = ridge_lambda))

cat("\n--- Ridge Regression ---\n")
cat("Optimal lambda:", round(ridge_lambda, 5), "\n")
cat("MAE: ",  round(mae(y_test, ridge_pred), 4), "\n")
cat("RMSE:", round(rmse(y_test, ridge_pred), 4), "\n")
cat("R²:  ",  round(cor(y_test, ridge_pred)^2, 4), "\n")
```

## 2.6 Model 3: Lasso Regression

```{r}
set.seed(42)
t_lasso <- system.time({
  set.seed(42)
  lasso_cv     <- cv.glmnet(X_train, y_train, alpha = 1, nfolds = 5)
  lasso_lambda <- lasso_cv$lambda.min
  lasso_model  <- glmnet(X_train, y_train, alpha = 1, lambda = lasso_lambda)
})["elapsed"]

lasso_pred <- as.vector(predict(lasso_model, newx = X_test, s = lasso_lambda))

# Show which features Lasso retained
lasso_coef <- coef(lasso_model, s = lasso_lambda)
cat("\n--- Lasso Regression ---\n")
cat("Optimal lambda:", round(lasso_lambda, 5), "\n")
cat("Non-zero coefficients:\n")
print(lasso_coef[lasso_coef[, 1] != 0, , drop = FALSE])
cat("MAE: ",  round(mae(y_test, lasso_pred), 4), "\n")
cat("RMSE:", round(rmse(y_test, lasso_pred), 4), "\n")
cat("R²:  ",  round(cor(y_test, lasso_pred)^2, 4), "\n")
```

## 2.7 Model 4: Random Forest

```{r}
train_rf <- data.frame(d_ln_ci = y_train, X_train)

t_rf <- system.time({
  train_rf <- data.frame(d_ln_ci = y_train, X_train)
  rf_model <- ranger(
    formula    = d_ln_ci ~ .,
    data       = train_rf,
    num.trees  = 500,
    importance = "permutation",
    seed       = 42
  )
})["elapsed"]

rf_pred <- predict(rf_model, data = as.data.frame(X_test))$predictions

cat("\n--- Random Forest ---\n")
cat("MAE: ",  round(mae(y_test, rf_pred), 4), "\n")
cat("RMSE:", round(rmse(y_test, rf_pred), 4), "\n")
cat("R²:  ",  round(cor(y_test, rf_pred)^2, 4), "\n")
```

## 2.8 Model 5: XGBoost

```{r}
# Cross-validation to find optimal number of rounds
t_xgb <- system.time({
  set.seed(42)
  dtrain <- xgb.DMatrix(data = X_train, label = y_train)
  xgb_params <- list(
    objective        = "reg:squarederror",
    eta              = 0.05,
    max_depth        = 4,
    subsample        = 0.8,
    colsample_bytree = 0.8
  )
  xgb_model <- xgb.train(
    params  = xgb_params,
    data    = dtrain,
    nrounds = 100,
    verbose = 0
  )
})["elapsed"]

dtest    <- xgb.DMatrix(data = X_test, label = y_test)
xgb_pred <- predict(xgb_model, dtest)

cat("\n--- XGBoost ---\n")
cat("MAE: ",  round(mae(y_test, xgb_pred), 4), "\n")
cat("RMSE:", round(rmse(y_test, xgb_pred), 4), "\n")
cat("R²:  ",  round(cor(y_test, xgb_pred)^2, 4), "\n")
```

## 2.9 Model Comparison Summary

```{r}
# Align naïve baseline to test set province-years for consistent comparison
# (naïve may lose one row per province due to double-lag)
naive_test_rows <- naive_pred %>%
  select(province, year, d_ln_ci, naive_pred)

# Collect all predictions on the common test set
results_all <- test %>%
  select(province, year, d_ln_ci) %>%
  mutate(
    ols_pred   = ols_pred,
    ridge_pred = ridge_pred,
    lasso_pred = lasso_pred,
    rf_pred    = rf_pred,
    xgb_pred   = xgb_pred
  )

# Build comparison table
model_names <- c("Naïve Persistence", "OLS", "Ridge", "Lasso",
                 "Random Forest", "XGBoost")

calc_metrics <- function(actual, pred) {
  c(
    MAE  = round(mae(actual, pred), 4),
    RMSE = round(rmse(actual, pred), 4),
    R2   = round(cor(actual, pred)^2, 4)
  )
}

metrics_table <- rbind(
  calc_metrics(naive_actual,      naive_hat),
  calc_metrics(y_test,            ols_pred),
  calc_metrics(y_test,            ridge_pred),
  calc_metrics(y_test,            lasso_pred),
  calc_metrics(y_test,            rf_pred),
  calc_metrics(y_test,            xgb_pred)
) %>%
  as.data.frame() %>%
  mutate(Model = model_names) %>%
  select(Model, MAE, RMSE, R2)

kable(metrics_table,
      caption = "Out-of-sample Performance Comparison (Test Set: 2018–2021)",
      align   = c("l", "r", "r", "r"))

# Save all predictions for use in Part 3
saveRDS(list(
  results_all  = results_all,
  rf_model     = rf_model,
  xgb_model    = xgb_model,
  lasso_model  = lasso_model,
  ridge_model  = ridge_model,
  ols_model    = ols_model,
  X_train      = X_train,
  X_test       = X_test,
  y_test       = y_test,
  lag_features = lag_features,
  metrics_table = metrics_table
), file = file.path(target_path, "model_objects.rds"))

cat("\nModel objects saved to model_objects.rds for use in Part 3.\n")
```

# Part 3: Model Interpretation and Comparative Analysis

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
library(ranger)
library(xgboost)
library(glmnet)
library(shapviz)
library(patchwork)

# Load model objects saved at end of Part 2
target_path <- "D:/yqy/硕士-mpp/第四学期/ds&climate change/project/data/target_variable"
obj <- readRDS(file.path(target_path, "model_objects.rds"))

results_all  <- obj$results_all
rf_model     <- obj$rf_model
xgb_model    <- obj$xgb_model
lasso_model  <- obj$lasso_model
ridge_model  <- obj$ridge_model
ols_model    <- obj$ols_model
X_train      <- obj$X_train
X_test       <- obj$X_test
y_test       <- obj$y_test
lag_features <- obj$lag_features
metrics_table <- obj$metrics_table

# Clean feature labels for plotting
feature_labels <- c(
  ln_gdp_percap_lag  = "Log GDP per Capita",
  urban_rate_lag     = "Urbanization Rate",
  sh_secondary_lag   = "Secondary Industry Share",
  sh_tertiary_lag    = "Tertiary Industry Share",
  coal_share_lag     = "Coal Share",
  PI_provincial_lag  = "Policy Intensity",
  ln_ci_lag          = "Log Carbon Intensity (lag)"
)
```

## 3.1 Prediction Performance: Actual vs. Predicted

```{r}
#| fig-width: 10
#| fig-height: 7

# Reshape to long format for faceted plot
pred_long <- results_all %>%
  pivot_longer(
    cols      = ends_with("_pred"),
    names_to  = "model",
    values_to = "predicted"
  ) %>%
  mutate(model = recode(model,
    ols_pred   = "OLS",
    ridge_pred = "Ridge",
    lasso_pred = "Lasso",
    rf_pred    = "Random Forest",
    xgb_pred   = "XGBoost"
  ))

ggplot(pred_long, aes(x = d_ln_ci, y = predicted)) +
  geom_point(alpha = 0.5, size = 1.5, color = "#2c7bb6") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
  facet_wrap(~ model, ncol = 3) +
  labs(
    title    = "Actual vs. Predicted Carbon Intensity Change (Test Set: 2018–2021)",
    subtitle = "Dashed red line = perfect prediction",
    x        = "Actual Δln(CI)",
    y        = "Predicted Δln(CI)"
  ) +
  theme_minimal(base_size = 12) +
  theme(strip.text = element_text(face = "bold"))
```

## 3.2 Prediction Error Distribution

```{r}
#| fig-width: 10
#| fig-height: 4

# Residuals for each model
residuals_long <- results_all %>%
  mutate(
    OLS           = d_ln_ci - ols_pred,
    Ridge         = d_ln_ci - ridge_pred,
    Lasso         = d_ln_ci - lasso_pred,
    `Random Forest` = d_ln_ci - rf_pred,
    XGBoost       = d_ln_ci - xgb_pred
  ) %>%
  select(province, year, OLS, Ridge, Lasso, `Random Forest`, XGBoost) %>%
  pivot_longer(-c(province, year), names_to = "model", values_to = "residual") %>%
  mutate(model = factor(model,
    levels = c("OLS", "Ridge", "Lasso", "Random Forest", "XGBoost")))

ggplot(residuals_long, aes(x = model, y = residual, fill = model)) +
  geom_boxplot(alpha = 0.7, outlier.size = 1) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey40") +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Distribution of Prediction Errors by Model",
    x     = NULL,
    y     = "Residual (Actual − Predicted)"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "none")
```

## 3.3 OLS and Lasso Coefficients

```{r}
#| fig-width: 9
#| fig-height: 4

# OLS coefficients with confidence intervals
ols_summary <- summary(ols_model)$coefficients %>%
  as.data.frame() %>%
  tibble::rownames_to_column("feature") %>%
  filter(feature != "(Intercept)") %>%
  rename(estimate = Estimate, se = `Std. Error`, p_value = `Pr(>|t|)`) %>%
  mutate(
    lower    = estimate - 1.96 * se,
    upper    = estimate + 1.96 * se,
    feature  = recode(feature, !!!feature_labels),
    sig      = ifelse(p_value < 0.05, "p < 0.05", "p ≥ 0.05")
  )

p_ols <- ggplot(ols_summary, aes(x = estimate, y = reorder(feature, estimate), color = sig)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0.3) +
  geom_point(size = 3) +
  scale_color_manual(values = c("p < 0.05" = "#d73027", "p ≥ 0.05" = "#91bfdb")) +
  labs(title = "OLS Coefficients (95% CI)", x = "Coefficient", y = NULL, color = NULL) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")

# Lasso coefficients
lasso_coef_df <- coef(lasso_model) %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column("feature") %>%
  rename(estimate = s0) %>%
  filter(feature != "(Intercept)", estimate != 0) %>%
  mutate(feature = recode(feature, !!!feature_labels))

p_lasso <- ggplot(lasso_coef_df, aes(x = estimate, y = reorder(feature, estimate))) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  geom_col(fill = "#4dac26", alpha = 0.8, width = 0.6) +
  labs(title = "Lasso Coefficients (Non-zero Only)", x = "Coefficient", y = NULL) +
  theme_minimal(base_size = 11)

p_ols + p_lasso
```

## 3.4 Feature Importance: Random Forest (Permutation)

```{r}
#| fig-width: 8
#| fig-height: 4

rf_imp <- rf_model$variable.importance %>%
  enframe(name = "feature", value = "importance") %>%
  mutate(
    feature = recode(feature, !!!feature_labels),
    importance_pct = importance / sum(importance) * 100
  ) %>%
  arrange(desc(importance))

ggplot(rf_imp, aes(x = importance_pct, y = reorder(feature, importance_pct))) +
  geom_col(fill = "lightblue", alpha = 0.85) +
  geom_text(aes(label = sprintf("%.1f%%", importance_pct)),
            hjust = -0.1, size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Random Forest: Permutation Feature Importance",
    x     = "Importance (%)",
    y     = NULL
  ) +
  theme_minimal(base_size = 12)
```

## 3.5 Feature Importance: XGBoost (Gain)

```{r}
#| fig-width: 8
#| fig-height: 4

xgb_imp <- xgb.importance(
  feature_names = lag_features,
  model         = xgb_model
) %>%
  as.data.frame() %>%
  mutate(
    Feature    = recode(Feature, !!!feature_labels),
    Gain_pct   = Gain / sum(Gain) * 100
  )

ggplot(xgb_imp, aes(x = Gain_pct, y = reorder(Feature, Gain_pct))) +
  geom_col(fill = "lightblue", alpha = 0.85) +
  geom_text(aes(label = sprintf("%.1f%%", Gain_pct)),
            hjust = -0.1, size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "XGBoost: Feature Importance (Gain)",
    x     = "Gain (%)",
    y     = NULL
  ) +
  theme_minimal(base_size = 12)
```

## 3.6 Feature Importance Comparison Across Models

```{r}
#| fig-width: 10
#| fig-height: 5

# Normalize all importances to 0–1 scale for cross-model comparison

# OLS: use absolute t-statistic as proxy for importance
ols_imp <- summary(ols_model)$coefficients %>%
  as.data.frame() %>%
  tibble::rownames_to_column("feature") %>%
  filter(feature != "(Intercept)") %>%
  transmute(
    feature    = recode(feature, !!!feature_labels),
    importance = abs(`t value`)
  ) %>%
  mutate(importance = importance / sum(importance), model = "OLS (|t-stat|)")

# Lasso: absolute coefficients
lasso_imp <- lasso_coef_df %>%
  transmute(
    feature    = feature,
    importance = abs(estimate)
  ) %>%
  # Fill in zeroed-out features
  right_join(
    tibble(feature = unname(feature_labels)),
    by = "feature"
  ) %>%
  mutate(
    importance = replace_na(importance, 0),
    importance = importance / sum(importance),
    model      = "Lasso (|coef|)"
  )

# RF: permutation importance (already normalized above)
rf_imp2 <- rf_imp %>%
  transmute(feature, importance = importance / sum(importance), model = "Random Forest")

# XGBoost: gain
xgb_imp2 <- xgb_imp %>%
  transmute(feature = Feature, importance = Gain / sum(Gain), model = "XGBoost")

importance_combined <- bind_rows(ols_imp, lasso_imp, rf_imp2, xgb_imp2) %>%
  mutate(model = factor(model,
    levels = c("OLS (|t-stat|)", "Lasso (|coef|)", "Random Forest", "XGBoost")))

ggplot(importance_combined,
       aes(x = importance, y = reorder(feature, importance), fill = model)) +
  geom_col(position = "dodge", alpha = 0.85) +
  scale_fill_brewer(palette = "Set1") +
  labs(
    title = "Feature Importance Comparison Across Models",
    x     = "Normalized Importance",
    y     = NULL,
    fill  = "Model"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
```

## 3.7 SHAP Values: Random Forest

```{r}
#| results: hide

library(kernelshap)

set.seed(42)
bg_idx <- sample(nrow(X_train), min(100, nrow(X_train)))

ks_rf <- kernelshap(
  rf_model,
  X        = X_test,
  bg_X     = X_train[bg_idx, ],
  pred_fun = function(model, X) predict(model, data = as.data.frame(X))$predictions
)
```

```{r}
#| fig-width: 9
#| fig-height: 5
shap_rf <- shapviz(ks_rf)

colnames(shap_rf$S) <- unname(feature_labels[colnames(shap_rf$S)])
colnames(shap_rf$X) <- unname(feature_labels[colnames(shap_rf$X)])

sv_importance(shap_rf, kind = "beeswarm") +
  labs(title = "SHAP Summary Plot — Random Forest (Test Set)") +
  theme_minimal(base_size = 12)
```

## 3.8 SHAP Values: XGBoost

```{r}
#| fig-width: 9
#| fig-height: 5

# XGBoost natively supports shapviz
shap_xgb <- shapviz(xgb_model, X_pred = X_test)

# Rename features for readability
colnames(shap_xgb$S) <- unname(feature_labels[lag_features])
colnames(shap_xgb$X) <- unname(feature_labels[lag_features])

sv_importance(shap_xgb, kind = "beeswarm") +
  labs(title = "SHAP Summary Plot — XGBoost (Test Set)") +
  theme_minimal(base_size = 12)
```

## 3.9 SHAP Dependence Plots: Top Features

```{r}
#| fig-width: 11
#| fig-height: 9

# Top 3 features by mean |SHAP| in XGBoost
top3 <- xgb_imp %>%
  slice_max(Gain, n = 3) %>%
  pull(Feature)

# Map back to original column names for shapviz
top3_orig <- names(feature_labels)[feature_labels %in% top3]

plots <- lapply(top3_orig, function(feat) {
  sv_dependence(shap_xgb, v = feature_labels[[feat]]) +
    theme_minimal(base_size = 11)
})

wrap_plots(plots, ncol = 1)
```

## 3.10 Summary and Key Findings

```{r}
# Reproduce the model comparison table from Part 2
kable(
  metrics_table,
  caption = "Out-of-sample Performance Comparison (Test Set: 2018–2021)",
  align   = c("l", "r", "r", "r")
)
```

```{r}
# Which model wins on each metric?
best_models <- metrics_table %>%
  filter(Model != "Naïve Persistence") %>%
  summarise(
    Best_MAE  = Model[which.min(MAE)],
    Best_RMSE = Model[which.min(RMSE)],
    Best_R2   = Model[which.max(R2)]
  )

cat("Best MAE: ",  best_models$Best_MAE,  "\n")
cat("Best RMSE:", best_models$Best_RMSE, "\n")
cat("Best R²:  ",  best_models$Best_R2,   "\n")

# Top feature by mean |SHAP| in XGBoost
top_feature <- xgb_imp %>% slice_max(Gain, n = 1) %>% pull(Feature)
cat("\nTop predictive feature (XGBoost Gain):", top_feature, "\n")

# Does policy intensity appear in Lasso-selected features?
policy_in_lasso <- "Policy Intensity" %in% lasso_coef_df$feature
cat("Policy Intensity retained by Lasso:", policy_in_lasso, "\n")
```

# Part 4: Calculating Carbon Emissions

```{r}
times_df <- data.frame(
  model   = c("Naive", "OLS", "Ridge", "Lasso", "Random Forest", "XGBoost"),
  seconds = c(t_naive, t_ols, t_ridge, t_lasso, t_rf, t_xgb)
)

carbon_path <- "D:/yqy/硕士-mpp/第四学期/ds&climate change/project/data/carbon_cost"

write.csv(times_df,
  file.path(carbon_path, "training_times.csv"),
  row.names = FALSE)
```

Carbon emissions were estimated using the \*\*CodeCarbon\*\* Python library. Training times for each model were first recorded in R using \`system.time()\` and exported to a CSV file. A separate Python script then read these times and used \`EmissionsTracker\` from CodeCarbon to estimate the carbon footprint of each model, by simulating the training duration on the same hardware.

CodeCarbon automatically identifies the local CPU specifications and regional electricity grid carbon intensity to compute emissions in gCO₂e. The results were saved to \`carbon_results.csv\` and loaded back into R for display.

```{r}
carbon <- read.csv(file.path(carbon_path, "carbon_results.csv"))

# Add total row
total_row <- data.frame(
  model      = "Total",
  seconds    = sum(carbon$seconds),
  emissions_g = sum(carbon$emissions_g)
)
carbon <- rbind(carbon, total_row)

kable(carbon,
  col.names = c("Model", "Training Time (s)", "Emissions (gCO₂e)"),
  caption   = "Estimated Carbon Cost of Model Training",
  digits    = 4)
```
